---
title: "Pearson Type IV curve fit to the frequency domain"
author: "Keyong"
date: "March 27, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's read in the data

```{r, message = F}
# Load the readxl library
library(readxl)

# Read in data as a data frame
fsp <- read_xlsx("17217_frequency_spectrum.xlsx")
fsp
```

Let's plot the data

```{r, message = FALSE, warning = FALSE}
# Load the dplyr and ggplot2 libraries
library(dplyr)
library(ggplot2)

# Plot the data as a line
ggplot(aes(freq, amp), data = fsp) +
    geom_line() + xlim(0, 50) + theme_classic() +
    ggtitle("Raw Data") +
    xlab("Frequency") + ylab("Amplitude")
```

To fit to a density function, we first need to calculate the **area under the curve**

```{r, message = FALSE, warning = FALSE}
# Load the DescTools library
library(DescTools)

# Calculate the area under the curve
area <- AUC(x = fsp$freq, y = fsp$amp)
area

```

We want to fit to a **Pearson Type IV density function**. Let's first try to use `geom_smooth()` (which uses the generalized additive model (gam) method):

```{r, message = FALSE, warning = FALSE}
# Load required libraries
# library(gsl)
library(PearsonDS)

# Define the formula for curve fitting
pearson4Curve <- amp ~ area * dpearsonIV(freq, m, nu, location, scale)

# Plot the data with a fitted curve
ggplot(aes(freq, amp), data = fsp) +
    geom_line() + xlim(0, 50) + theme_classic() +
    geom_smooth(formula = pearson4Curve) +
    ggtitle("Test fit to the frequency spectrum with GAM") +
    xlab("Frequency") + ylab("Amplitude")
```

Since it didn't work very well, let's first use trial and error to find a Pearson Type IV density curve that approximately matches our data:

```{r, message = FALSE, warning = FALSE}
# Define starting parameters
m0 <- 1
nu0 <- -6
location0 <- 0.5
scale0 <- 0.5

# Add a column for the predicted amplitude values
fsp <- 
    fsp %>% 
    mutate(ampPredicted = area * 
            dpearsonIV(freq, m = m0, nu = nu0, location = location0, scale = scale0))

# Plot the data with the Pearson Type IV curve
ggplot(aes(freq, amp), data = fsp) +
    geom_line(color = 'Blue') + xlim(0, 50) + theme_classic() +
    geom_line(aes(freq, ampPredicted), color = 'Red') +
    ggtitle("Approximate fit to the frequency spectrum by trial and error") +
    xlab("Frequency") + ylab("Amplitude")
```

Now we have initial guesses for the 4 parameters, let's use **nonlinear least squares** method to get a better fit to the data:

```{r, message = FALSE, warning = FALSE}
# Use nonlinear least squares to do curve fitting
model <- 
    nls(formula = pearson4Curve, data = fsp, 
        start = list(m = m0, nu = nu0, location = location0, scale = scale0))
model
```

Use the **broom** package to tidy the results

```{r, message = FALSE, warning = FALSE}
# Import the broom library
library(broom)

# Tidy the model
results <- tidy(model)
results
```


Finally, plot the data with the refined Pearson Type IV model

```{r, message = FALSE, warning = FALSE}
# Add a column for the predicted amplitude values
fsp <- 
    fsp %>% 
    mutate(ampPredicted = area * 
            dpearsonIV(freq, m = results$estimate[1], nu = results$estimate[2], 
                       location = results$estimate[3], scale = results$estimate[4]))

# Plot the data with the Pearson Type IV curve
ggplot(aes(freq, amp), data = fsp) +
    geom_line(color = 'Blue') + xlim(0, 50) + theme_classic() +
    geom_line(aes(freq, ampPredicted), color = 'Red') +
    ggtitle("Best fit to the frequency spectrum by nls") +
    xlab("Frequency") + ylab("Amplitude")
```

```{r}
# Find the maximum of the fit
maxAmpPredicted <- max(fsp$ampPredicted)
maxAmpPredicted

# Print the row wihthe maximum amplitude
fsp %>% 
  filter(ampPredicted == maxAmpPredicted)

```

